The Private School Network: Recruiting Visits to Private High Schools by Public and Private Universities

### 2024-03-31 manuscript/manuscript_rihe_r_r.Rmd
#18 remove files

Scholarship on nonresident enrollment by public research universities has developed in isolation from scholarship on linkages between private high schools and selective private universities. We argue that these literatures are part of a broader story about the competition for students from affluent schools and communities. This manuscript analyzes off-campus recruiting visits to private high schools made by a convenience sample of 15 public research universities and 14 selective private universities. An off-campus recruiting visit indicates a social relationship between a school and a university. Therefore, we utilize social network methods to examine the recruiting networks of public and private universities. With respect to scale (research question (RQ) 1), universities in our sample made a disproportionate number of visits to private high schools. With respect to overlap (RQ2), simple network analyses and community detection methods reveal substantial overlap in the recruiting networks of public and private universities. RQ3 assesses the characteristics of visited schools. Both public and private universities tended to visit private schools in their home geographic region and also in the South, where private school enrollment growth has been strongest. Visited private schools enroll a much larger share of white students than visited public schools. Surprisingly, several public research universities visited sectarian private high schools at a rate similar to sectarian private universities.

@@

During the 2017 calendar year, we tracked off-campus recruiting visits made by 14 selective private universities, 12 selective private liberal arts colleges, and 15 public research universities.^[These institutions are: Northwestern, Notre Dame, Emory, Tufts, Boston College, Tulane, Case Western Reserve, Villanova, Southern Methodist University, Baylor, University of Denver, Texas Christian University, Stevens Institute of Technology, Marquette, Williams, Swarthmore, Middlebury, Smith, Harvey Mudd, Colorado College, Macalester, Scripps, Oberlin, Occidental, Sewanee, Connecticut College, UC-Berkeley, UC-San Diego, University of Georgia, University of Pittsburgh, Rutgers, University of Massachusetts-Amherst, UC-Riverside, SUNY Stony Brook, University of Colorado-Boulder, University of South Carolina, University of Kansas, University of Nebraska-Lincoln, University of Alabama, University of Cincinnati, and University of Arkansas.] From September 6th through October 17th, Harpeth Hall received recruiting visits from 11 private universities in our sample (Northwestern, Boston College, Tufts, Villanova, Case Western Reserve, University of Denver, Emory, Tulane, Baylor, Texas Christian University, Southern Methodist University), and three private liberal arts colleges (Harvey Mudd, Scripps, and Sewanee). Previous scholarship observes strong connections between private schools and selective private universities [@RN3519; @RN4407]. However, Harpeth Hall also received visits from six out-of-state public research universities, the University of Cincinnati, University of South Carolina, the University of Alabama, University of Arkansas, University of Georgia, and CU Boulder.

@@

Christ Presbyterian Academy, located in the Forest Hills suburb southwest of Nashville, received an "A" grade from Niche, meaning it is not ranked in the top 1,000 U.S. private schools. In 2017, it received visits from six private universities in our sample (including the three Christian universities, Baylor, Texas Christian University, and Southern Methodist University). It also received visits from the University of Cincinnati, the University of South Carolina, the University of Alabama, University of Arkansas, and CU Boulder. Christ Presbyterian Academy was founded in 1985 and 94% of its students are white.

This paper analyzes the social networks revealed by recruiting visits from universities to private high schools. Theories of organizational behavior suggest that university recruiting behaviors -- those that require significant expenditure -- are indicators of organizational enrollment priorities. Public-facing organizations are beset by multiple, competing pressures from the external environment, but organizations have difficulty pursuing multiple goals [@RN531; @RN959]. Thus, organizations respond by symbolically adopting some goals and substantively adopting others [@RN513]. Under symbolic adoption, organizations adopt policies and structures that send signals to external constituents but do not substantively affect resource allocation inside the "technical level" of the organization.^[Contingency theory states that organizations consist of "technical levels" responsible for processing inputs and producing outputs and "managerial levels" responsible for directing the technical level and tending to demands from the external environment [@RN531]. The technical level achieves effectiveness by focusing on a small, stable set of goals.] Under substantive adoption, organizations direct resources in the technical level towards achieving the goal. Off-campus recruiting visits to local high schools and college fairs was the top-ranked recruiting expenditure item for both public and private universities, according to @RN4664. Therefore, analyses of which schools a university visits -- and which schools they ignore -- can yield insights about university enrollment goals.

We situate this study within scholarship on nonresident enrollment at public universities. @RN4758 analyzed recruiting visits to public high schools made by 15 public research universities in 2017. The authors found that 12 of the 15 universities made more visits to out-of-state schools than in-state schools. These out-of-state visits focused on affluent, predominantly white public schools. Set against the scholarship about declining state appropriations and growing nonresident enrollment [@RN3753; @RN3685], these results suggest that many public research universities prioritize enrollment from wealthy out-of-state students who can afford to pay nonresident tuition prices.

@@

We argue that these two literatures have become connected because public research universities increasingly target out-of-state private schools, schools that have been the domain of selective private universities. Simply put, selective private universities and public research universities are increasingly competing for students from the same set of affluent households.

@@

Both private and public universities in our sample made a disproportionate number of visits to private high schools. For public research universities, visited private schools were much more likely to be out-of-state than visited public schools. Network analyses revealed substantial overlap in the recruiting networks of universities in our sample. For several public universities (e.g., University of Alabama, University of South Carolina), their recruiting networks had more similarity to private universities (e.g., Southern Methodist University, Texas Christian University) than they did to other public universities in our sample. Results suggest that recruiting networks are substantially a function of geographic location. All universities visited a large number of private schools in their home region. However, most universities in our sample also visited a large number of private schools in the South, the region where private school enrollment has grown the most over the past three decades [@RN4504]. The private schools visited by public and private universities in our sample enrolled a much higher share of white students than the public schools visited by these universities. Surprisingly, several public universities visited sectarian private high schools at a rate similar to sectarian private universities.

The results yield practical implications for public and private universities concerned about racial and socioeconomic diversity. Following the COVID-19 Pandemic, off-campus recruiting visits have bounced back because they remain a fundamental means of maintaining relationships with high school guidance counselors [@RN4896; @RN4895; @eab_nacac_counselor_2023]. Prior research finds that students are sensitive to recruiting interventions, including off-campus recruiting visits [@RN4324; @RN4402]. Therefore, which high schools a university chooses to visit affects the subsequent enrollment composition of the university. In the aftermath of the Supreme Court decision on race-conscious admissions, public and private universities are searching for ways to increase racial diversity without running afoul of the law. The @RN4954[p. 9] recommends that universities expand "outreach to high schools with substantial populations of low-income students and students of color." Most public and private universities in our sample visited a disproportionate number of private schools and these private schools tended to enroll mostly white students. We recommend that universities concerned about racial diversity should conduct an internal equity audit of their recruiting practices, for example, comparing the racial and socioeconomic characteristics of schools they visit versus schools they do not visit.

@@

Other studies of nonresident enrollment examine university enrollment composition. Using data from NPSAS 2011-12, @RN3685 show that compared to resident students at public research universities, nonresident students generate higher net tuition revenue, are more affluent, are less likely to be Pell recipients, and are less likely to be Black or Latinx. Institution-level panel models reveal that growth in the share of nonresident students was associated with a decline in the share of Pell recipients and a decline in the share of Black and Latinx students. @RN4290 examine whether growth in the number of nonresident students causes resident enrollment to decline. They find that nonresident enrollment "crowds out" resident enrollment at prestigious public research universities, but not at less prestigious public research universities.

@@

Research question 1 states, "How does the scale of visits to private high schools vary for the public research universities and selective private universities in our sample?" We answer this research question by comparing degree centrality across universities, both within university type (public vs. private) and between university types. In order to discern the relative emphasis on private schools by each university, we also compare the number of recruiting visits to private schools to the number of visits to public schools.

@@

**RQ3: Ego networks and homophily**. Because the data structure of networks is complicated, it is often illuminating to analyze "ego networks" rather than the entire social network. An "ego" is a single "focal" node (e.g., The University of Notre Dame). An ego network consists "of a focal node ('ego') and the nodes to whom ego is directly connected to (these are called 'alters') plus the ties, if any, among the alters" [@RN4953]. An ego network of order 1 consists of the alters that are directly tied to the ego. The order 1 ego network for the University of Notre Dame consists of all private high schools that received at least one visit from Notre Dame. The order 2 ego network consists of all private schools that were visited by Notre Dame and all universities that visited the private high schools visited by Notre Dame. Figure \@ref(fig:nd-religion) shows the order 2 ego network for Notre Dame, with vertices colored by religious affiliation. Universities that are closer to Notre Dame visit more private high schools in common.

@@

Our analyses focus on the following private school characteristics: geographic region, religious affiliation, academic reputation, racial composition, and enrollment size. Decisions about which private school characteristics to analyze were informed by logical argument and by prior research [e.g., @RN4724; @RN4504; @RN4321; @RN3519; @RN4588]. We analyze geographic region because, for most colleges and universities, enrollment demand is likely to be stronger at high schools in the same geographic region, but nationally prestigious colleges and universities may experience high demand throughout the country. We analyze religious affiliation because private colleges are more likely to have a relationship with private high schools with a similar religious affiliation. Although public universities are formally secular, we are interested in whether particular public universities tend to target religiously affiliated private high schools. We analyze academic reputation based on the idea that more highly ranked colleges are likely to target highly ranked private high schools. Racial composition is of substantive interest because the racial composition of visited high schools has consequences for the racial composition of colleges and universities. We would also like to analyze tuition price and household income, but these data are not available for private high schools.

@@

With respect to research question 1, the scale of visits to private high schools by public research universities ranged from a low of `r events_df %>% filter(univ_id == '110671', event_type %in% c('priv_hs')) %>% count() %>% as.numeric()` visits by UC-Riverside to a high of `r events_df %>% filter(univ_id == '100751', event_type %in% c('priv_hs')) %>% count() %>% as.numeric()` visits by the University of Alabama (which accounted for `r round((events_df %>% filter(univ_id == '100751', event_type %in% c('priv_hs')) %>% count() %>% as.numeric())/(events_df %>% filter(classification == 'public_research', event_type %in% c('priv_hs')) %>% count() %>% as.numeric())*100, digits=1)`% of all visits to private schools by public research universities in our sample), with the median university, The University of Pittsburgh, making `r events_df %>% filter(classification == 'public_research', event_type %in% c('priv_hs')) %>% count(univ_abbrev) %>% arrange(n) %>% summarize(median = median(n)) %>% as.numeric()` visits.  Across all public research universities, `r round((events_df %>% filter(classification == 'public_research', event_type %in% c('priv_hs'),univ_state != event_state) %>% count() %>% as.numeric())/(events_df %>% filter(classification == 'public_research', event_type %in% c('priv_hs')) %>% count() %>% as.numeric())*100, digits=1)`% of private high school visits were to out-of-state schools while `r round((events_df %>% filter(classification == 'public_research', event_type %in% c('pub_hs'),univ_state != event_state) %>% count() %>% as.numeric())/(events_df %>% filter(classification == 'public_research', event_type %in% c('pub_hs')) %>% count() %>% as.numeric())*100, digits=1)`% of public high school visits were to out-of-state schools. This result is consistent with the notion that public research universities placed more value on out-of-state private high school students than in-state private high school students because out-of-state students generate substantially more tuition revenue.

For selective private universities, the number of visits to private high schools ranged from a low of `r events_df %>% filter(univ_id == '186867', event_type %in% c('priv_hs')) %>% count() %>% as.numeric()` by Stevens Institute of Technology to a high of `r events_df %>% filter(univ_id == '152080', event_type %in% c('priv_hs')) %>% count() %>% as.numeric()` by Notre Dame, with the median universities -- Boston College and Northwestern University -- visiting `r events_df %>% filter(univ_id == '164924', event_type %in% c('priv_hs')) %>% count() %>% as.numeric()` and `r events_df %>% filter(univ_id == '147767', event_type %in% c('priv_hs')) %>% count() %>% as.numeric()` private high schools, respectively. Thus, for the universities in our sample, the scale of visits to private high schools tended to be higher for selective private universities than public research universities. That said, several public research universities in our sample -- CU-Boulder (`r events_df %>% filter(univ_id == '126614', event_type %in% c('priv_hs')) %>% count() %>% as.numeric()` visits), University of South Carolina (`r events_df %>% filter(univ_id == '218663', event_type %in% c('priv_hs')) %>% count() %>% as.numeric()` visits), and University of Alabama (`r events_df %>% filter(univ_id == '100751', event_type %in% c('priv_hs')) %>% count() %>% as.numeric()` visits) -- made a relatively high number of private high school visits even when compared to the selective private universities in our sample.

The relative scale of visits to private high schools -- defined as private high school visits as a percentage of all high school visits -- also tended to be substantially higher at selective private universities than at public research universities.  Across all selective private universities in our sample, `r round((events_df %>% filter(classification == 'private_national', event_type %in% c('priv_hs')) %>% count() %>% as.numeric())/(events_df %>% filter(classification == 'private_national', event_type %in% c('priv_hs','pub_hs')) %>% count() %>% as.numeric())*100, digits=1)`% of high school visits were to private high schools, compared to `r round((events_df %>% filter(classification == 'public_research', event_type %in% c('priv_hs')) %>% count() %>% as.numeric())/(events_df %>% filter(classification == 'public_research', event_type %in% c('priv_hs','pub_hs')) %>% count() %>% as.numeric())*100, digits=1)`% for public research universities. Although, the relative emphasis on private high schools was slightly closer for out-of-state visits; across all selective private universities, `r round((events_df %>% filter(classification == 'private_national', event_type %in% c('priv_hs'),univ_state != event_state) %>% count() %>% as.numeric())/(events_df %>% filter(classification == 'private_national', event_type %in% c('priv_hs','pub_hs'),univ_state != event_state) %>% count() %>% as.numeric())*100, digits=1)`% of out-of-state high school visits were to private schools, compared to `r round((events_df %>% filter(classification == 'public_research', event_type %in% c('priv_hs'),univ_state != event_state) %>% count() %>% as.numeric())/(events_df %>% filter(classification == 'public_research', event_type %in% c('priv_hs','pub_hs'),univ_state != event_state) %>% count() %>% as.numeric())*100, digits=1)`% for public research universities.

Even though selective private universities tend to emphasize visits to private high schools more than public research universities, both groups visited private high schools at a much higher rate than public high schools. Figure \@ref(fig:actual-proportional-pubu) (public universities) and Figure \@ref(fig:actual-proportional-privu) (private universities) show the actual number of private school visits made compared to the hypothetical number of private school visits if each public and private high school in the nation had an equal probability of receiving a visit.^[When calculating the hypothetical proportional number of visits for each university, we only consider high schools in states that received at least one visit to a public or private high school from that university. Our rationale for this decision is that it is unhelpful to include schools in states that the university ignored entirely.] Figure \@ref(fig:actual-proportional-pubu) shows that the actual number of private school visits exceeded the proportional number of private school visits for 13 of the 15 public research universities in our sample, the exceptions being UC-Riverside and SUNY Stony Brook. For UC-Berkeley, private school visits represented 27% of all high school visits but would have represented 19% had UC-Berkeley visited public and private schools at the same rate. For the University of Georgia, which represented the extreme for public universities, private school visits represented 37% of all high school visits but would have represented 19% had the university visited public and private schools at the same rate. Figure \@ref(fig:actual-proportional-privu) shows that this disparity between actual versus proportional visits was much greater for private universities than public universities. Catholic universities made the most disproportionate number of private high school visits. For Notre Dame, private school visits represented 63% of actual high school visits but would have represented 18% of proportional visits.

@@

Research question 2 asks, "To what extent do public research universities and selective private universities visit overlapping sets of private high schools?" Answering this question requires social network analysis techniques. As described in the methods section, we began by constructing a two-mode social network of visits to private high schools (mode 1) by universities (mode 2). Each university and each private high school is a "vertex." A visit from a university to a school is an "edge." Because large two-mode networks like our school-university network are difficult to visualize, two-mode networks are often analyzed as one-mode networks [@RN2457;@RN4663]. We transformed our two-mode school-university network into a one-mode university network in which each university is a vertex and two universities share an edge if both universities visited at least one high school in common. The *weight* of an edge indicates the number of high schools that both universities visited. For example, Appendix Table \@ref(tab:pubu-count-matrix) shows the matrix of weights for the set of public research universities. We created three separate one-mode university networks: a public university network; a private university network; a public and private university network.

@@

Table \@ref(tab:pubu-percent-matrix) provides a tabular representation of the extent to which public universities visited overlapping sets of private high schools. For each public university (column), each cell of Table \@ref(tab:pubu-percent-matrix) shows the number of visited high schools in common with the row university as a percent of the total number of high schools visited by the column university. This indicates the extent to which public universities were competing for students at the same set of private schools. For example, column 2 shows that of the 396 different private high schools visited by the University of South Carolina, the University of Alabama visited 71.7% of these high schools and CU-Boulder visited 39.1% of these high schools. Appendix Table \@ref(tab:pubu-count-matrix) shows counts rather than percentages.

@@

Next, we examine overlap in a network that contains the 15 public universities and the 14 private universities. Table \@ref(tab:privu-percent-visited-by-pubu-matrix) shows, for each private university, the percentage of schools that were also visited by each public university. This table shows the extent to which public universities were competing at private schools targeted by private universities. For example, of the 625 private schools visited by Notre Dame, the University of Alabama visited 52.2% of these schools. Similarly, CU-Boulder visited 46.4% of the 435 schools that TCU visited and 42.4% of the 377 schools that Northwestern visited. The University of Arkansas competed heavily at schools visited by Baylor; Rutgers and UMass-Amherst competed heavily at schools visited by Stevens Institute of Technology; and UC-Berkeley competed at schools visited by Northwestern, Tufts, and Emory.

@@

Universities located near the center of Figure \@ref(fig:plot-1mode-u) tended to be universities with the highest degree centrality (i.e., those that visited the most schools). The top nine universities in terms of degree centrality are (number of unique private high schools visited in parentheses): University of Alabama (759), Notre Dame (625), Villanova (563), SMU (550), TCU (435), Tulane (430), University of South Carolina (396), Northwestern (377), and CU-Boulder (362). Notably, while the majority of universities located near the center of Figure \@ref(fig:plot-1mode-u) are private, three are public. The most peripheral universities -- UC-Riverside,  SUNY Stony Brook, and University of Nebraska-Lincoln -- are all public.

>>>>

Scholarship on nonresident enrollment by public research universities has developed in isolation from scholarship on linkages between private high schools and selective private universities. We argue that these literatures are part of a broader story about the competition for students from affluent schools and communities. This manuscript analyzes off-campus recruiting visits to private high schools made by a convenience sample of 15 public research universities and 14 selective private universities. An off-campus recruiting visit indicates a social relationship between a school and a university. Therefore, we utilize social network methods to examine the recruiting networks of public and private universities. With respect to scale (research question 1), universities in our sample made a disproportionate number of visits to private high schools. With respect to overlap (RQ2), simple network analyses and community detection methods reveal substantial overlap in the recruiting networks of public and private universities. RQ3 assesses the characteristics of visited schools. Both public and private universities tended to visit private high schools in their home geographic region and also in the South, where private school enrollment growth has been strongest. Visited private high schools enroll a much larger share of white students than visited public high schools. Surprisingly, several public research universities visited sectarian private high schools at a rate similar to sectarian private universities.

@@

During the 2017 calendar year, we tracked off-campus recruiting visits made by 14 selective private universities, 12 selective private liberal arts colleges, and 15 public research universities.^[These institutions are: Northwestern University, University of Notre Dame, Emory University, Tufts University, Boston College, Tulane University, Case Western Reserve University, Villanova University, Southern Methodist University, Baylor University, University of Denver, Texas Christian University, Stevens Institute of Technology, Marquette University, Williams College, Swarthmore College, Middlebury College, Smith College, Harvey Mudd College, Colorado College, Macalester College, Scripps College, Oberlin College, Occidental College, Sewanee, Connecticut College, UC-Berkeley, UC-San Diego, University of Georgia, University of Pittsburgh, Rutgers University, University of Massachusetts-Amherst, UC-Riverside, SUNY Stony Brook, University of Colorado-Boulder, University of South Carolina, University of Kansas, University of Nebraska-Lincoln, University of Alabama, University of Cincinnati, and University of Arkansas.] From September 6th through October 17th, Harpeth Hall received recruiting visits from 11 private universities in our sample (Northwestern, Boston College, Tufts, Villanova, Case Western Reserve, University of Denver, Emory, Tulane, Baylor, TCU, Southern Methodist University), and three private liberal arts colleges (Harvey Mudd, Scripps, and Sewanee). Previous scholarship observes strong connections between private schools and selective private universities [@RN3519; @RN4407]. However, Harpeth Hall also received visits from six out-of-state public research universities, the University of Cincinnati, University of South Carolina, the University of Alabama, University of Arkansas, University of Georgia, and CU Boulder.

@@

Christ Presbyterian Academy, located in the Forest Hills suburb southwest of Nashville, received an "A" grade from Niche, meaning it is not ranked in the top 1,000 U.S. private schools. In 2017, it received visits from six private universities in our sample (including the three Christian universities, Baylor, TCU, and Southern Methodist University). It also received visits from the University of Cincinnati, the University of South Carolina, the University of Alabama, University of Arkansas, and CU Boulder. Christ Presbyterian Academy was founded in 1985 and 94% of its students are white.

This paper analyzes the social networks revealed by recruiting visits from universities to private high schools. Theories of organizational behavior suggest that university recruiting behaviors -- those that require significant expenditure -- are indicators of organizational enrollment priorities. Public-facing organizations are beset by multiple, competing pressures from the external environment, but organizations have difficulty pursuing multiple goals [@RN531; @RN959]. Thus, organizations respond by symbolically adopting some goals and substantively adopting others [@RN513]. Under symbolic adoption, organizations adopt policies and structures that send signals to external constituents but do not substantively affect resource allocation inside the "technical level" of the organization.^[Contingency theory states that organizations consist of "technical levels" responsible for processing inputs and producing outputs and "managerial levels" responsible for directing the technical level and tending to demands from the external environment [@RN531]. The technical level achieves effectiveness by focusing on a small, stable set of goals.] Under substantive adoption, organizations direct resources in the technical level towards achieving the goal. Off-campus recruiting visits to local high schools and college fairs was the top-ranked recruiting expenditure item for both public and private universities, according to @RN4664. Therefore, analyses of which high schools a university visits -- and which high schools they ignore -- can yield insights about university enrollment goals.

We situate this study within scholarship on nonresident enrollment at public universities. @RN4758 analyzed recruiting visits to public high schools made by 15 public research universities in 2017. The authors found that 12 of the 15 universities made more visits to out-of-state schools than in-state schools. These out-of-state visits focused on affluent, predominantly white public high schools. Set against the scholarship about declining state appropriations and growing nonresident enrollment [@RN3753; @RN3685], these results suggest that many public research universities prioritize enrollment from wealthy out-of-state students who can afford to pay nonresident tuition prices.

@@

We argue that these two literatures have become connected because public research universities increasingly target out-of-state private high schools, schools that have been the domain of selective private universities. Simply put, selective private universities and public research universities are increasingly competing for students from the same set of affluent households.

@@

Both private and public universities in our sample made a disproportionate number of visits to private high schools. For public research universities, visited private high schools were much more likely to be out-of-state than visited public high schools. Network analyses revealed substantial overlap in the recruiting networks of universities in our sample. For several public universities (e.g., University of Alabama, University of South Carolina), their recruiting networks had more similarity to private universities (e.g., Southern Methodist University, TCU) than they did to other public universities in our sample. Results suggest that recruiting networks are substantially a function of geographic location. All universities visited a large number of private high schools in their home region. However, most universities in our sample also visited a large number of private high schools in the South, the region where private school enrollment has grown the most over the past three decades [@RN4504]. The private high schools visited by public and private universities in our sample enrolled a much higher share of white students than the public high schools visited by these universities. Surprisingly, several public universities visited sectarian private high schools at a rate similar to sectarian private universities.

The results yield practical implications for public and private universities concerned about racial and socioeconomic diversity. Following the COVID-19 Pandemic, off-campus recruiting visits have bounced back because they remain a fundamental means of maintaining relationships with high school guidance counselors [@RN4896; @RN4895; @eab_nacac_counselor_2023]. Prior research finds that students are sensitive to recruiting interventions, including off-campus recruiting visits [@RN4324; @RN4402]. Therefore, which high schools a university chooses to visit affects the subsequent enrollment composition of the university. In the aftermath of the Supreme Court decision on race-conscious admissions, public and private universities are searching for ways to increase racial diversity without running afoul of the law. The @RN4954[p. 9] recommends that universities expand "outreach to high schools with substantial populations of low-income students and students of color." Most public and private universities in our sample visited a disproportionate number of private high schools and these private schools tended to enroll mostly white students. We recommend that universities concerned about racial diversity should conduct an internal equity audit of their recruiting practices, for example, comparing the racial and socioeconomic characteristics of schools they visit versus schools they do not visit.

@@

Other studies of nonresident enrollment examine university enrollment composition. Using data from NPSAS 2011-12, @RN3685 show that compared to resident students at public research universities, nonresident students generate higher net tuition revenue, are more affluent, are less likely to be Pell Grant recipients, and are less likely to be Black or Latinx. Institution-level panel models reveal that growth in the share of nonresident students was associated with a decline in the share of Pell Grant recipients and a decline in the share of Black and Latinx students. @RN4290 examine whether growth in the number of nonresident students causes resident enrollment to decline. They find that nonresident enrollment "crowds out" resident enrollment at prestigious public research universities, but not at less prestigious public research universities.

@@

Research question 1 states, "How does the scale of visits to private high schools vary for the public research universities and selective private universities in our sample?" We answer this research question by comparing degree centrality across universities, both within university type (public vs. private) and between university types. In order to discern the relative emphasis on private high schools by each university, we also compare the number of recruiting visits to private high schools to the number of visits to public high schools.

@@

**RQ3: Ego networks and homophily**. Because the data structure of networks is complicated, it is often illuminating to analyze "ego networks" rather than the entire social network. An "ego" is a single "focal" node (e.g., Notre Dame). An ego network consists "of a focal node ('ego') and the nodes to whom ego is directly connected to (these are called 'alters') plus the ties, if any, among the alters" [@RN4953]. An ego network of order 1 consists of the alters that are directly tied to the ego. The order 1 ego network for Notre Dame consists of all private high schools that received at least one visit from Notre Dame. The order 2 ego network consists of all private schools that were visited by Notre Dame and all universities that visited the private high schools visited by Notre Dame. Figure \@ref(fig:nd-religion) shows the order 2 ego network for Notre Dame, with vertices colored by religious affiliation. Universities that are closer to Notre Dame visit more private high schools in common.

@@

Our analyses focus on the following private high school characteristics: geographic region, religious affiliation, academic reputation, racial composition, and enrollment size. Decisions about which private high school characteristics to analyze were informed by logical argument and by prior research [e.g., @RN4724; @RN4504; @RN4321; @RN3519; @RN4588]. We analyze geographic region because, for most colleges and universities, enrollment demand is likely to be stronger at high schools in the same geographic region, but nationally prestigious colleges and universities may experience high demand throughout the country. We analyze religious affiliation because private colleges are more likely to have a relationship with private high schools with a similar religious affiliation. Although public universities are formally secular, we are interested in whether particular public universities tend to target religiously affiliated private high schools. We analyze academic reputation based on the idea that more highly ranked colleges are likely to target highly ranked private high schools. Racial composition is of substantive interest because the racial composition of visited high schools has consequences for the racial composition of colleges and universities. We would also like to analyze tuition price and household income, but these data are not available for private high schools.

@@

With respect to research question 1, the scale of visits to private high schools by public research universities ranged from a low of `r events_df %>% filter(univ_id == '110671', event_type %in% c('priv_hs')) %>% count() %>% as.numeric()` visits by UC-Riverside to a high of `r events_df %>% filter(univ_id == '100751', event_type %in% c('priv_hs')) %>% count() %>% as.numeric()` visits by the University of Alabama (which accounted for `r round((events_df %>% filter(univ_id == '100751', event_type %in% c('priv_hs')) %>% count() %>% as.numeric())/(events_df %>% filter(classification == 'public_research', event_type %in% c('priv_hs')) %>% count() %>% as.numeric())*100, digits=1)`% of all visits to private high schools by public research universities in our sample), with the median university, The University of Pittsburgh, making `r events_df %>% filter(classification == 'public_research', event_type %in% c('priv_hs')) %>% count(univ_abbrev) %>% arrange(n) %>% summarize(median = median(n)) %>% as.numeric()` visits.  Across all public research universities, `r round((events_df %>% filter(classification == 'public_research', event_type %in% c('priv_hs'),univ_state != event_state) %>% count() %>% as.numeric())/(events_df %>% filter(classification == 'public_research', event_type %in% c('priv_hs')) %>% count() %>% as.numeric())*100, digits=1)`% of private high school visits were to out-of-state high schools while `r round((events_df %>% filter(classification == 'public_research', event_type %in% c('pub_hs'),univ_state != event_state) %>% count() %>% as.numeric())/(events_df %>% filter(classification == 'public_research', event_type %in% c('pub_hs')) %>% count() %>% as.numeric())*100, digits=1)`% of public high school visits were to out-of-state high schools. This result is consistent with the notion that public research universities placed more value on out-of-state private high school students than in-state private high school students because out-of-state students generate substantially more tuition revenue.

For selective private universities, the number of visits to private high schools ranged from a low of `r events_df %>% filter(univ_id == '186867', event_type %in% c('priv_hs')) %>% count() %>% as.numeric()` by Stevens Institute of Technology to a high of `r events_df %>% filter(univ_id == '152080', event_type %in% c('priv_hs')) %>% count() %>% as.numeric()` by Notre Dame, with the median universities -- Boston College and Northwestern University -- visiting `r events_df %>% filter(univ_id == '164924', event_type %in% c('priv_hs')) %>% count() %>% as.numeric()` and `r events_df %>% filter(univ_id == '147767', event_type %in% c('priv_hs')) %>% count() %>% as.numeric()` private high schools, respectively. Thus, for the universities in our sample, the scale of visits to private high schools tended to be higher for selective private universities than public research universities. That said, several public research universities in our sample -- CU Boulder (`r events_df %>% filter(univ_id == '126614', event_type %in% c('priv_hs')) %>% count() %>% as.numeric()` visits), University of South Carolina (`r events_df %>% filter(univ_id == '218663', event_type %in% c('priv_hs')) %>% count() %>% as.numeric()` visits), and University of Alabama (`r events_df %>% filter(univ_id == '100751', event_type %in% c('priv_hs')) %>% count() %>% as.numeric()` visits) -- made a relatively high number of private high school visits even when compared to the selective private universities in our sample.

The relative scale of visits to private high schools -- defined as private high school visits as a percentage of all high school visits -- also tended to be substantially higher at selective private universities than at public research universities.  Across all selective private universities in our sample, `r round((events_df %>% filter(classification == 'private_national', event_type %in% c('priv_hs')) %>% count() %>% as.numeric())/(events_df %>% filter(classification == 'private_national', event_type %in% c('priv_hs','pub_hs')) %>% count() %>% as.numeric())*100, digits=1)`% of high school visits were to private high schools, compared to `r round((events_df %>% filter(classification == 'public_research', event_type %in% c('priv_hs')) %>% count() %>% as.numeric())/(events_df %>% filter(classification == 'public_research', event_type %in% c('priv_hs','pub_hs')) %>% count() %>% as.numeric())*100, digits=1)`% for public research universities. Although, the relative emphasis on private high schools was slightly closer for out-of-state visits; across all selective private universities, `r round((events_df %>% filter(classification == 'private_national', event_type %in% c('priv_hs'),univ_state != event_state) %>% count() %>% as.numeric())/(events_df %>% filter(classification == 'private_national', event_type %in% c('priv_hs','pub_hs'),univ_state != event_state) %>% count() %>% as.numeric())*100, digits=1)`% of out-of-state high school visits were to private high schools, compared to `r round((events_df %>% filter(classification == 'public_research', event_type %in% c('priv_hs'),univ_state != event_state) %>% count() %>% as.numeric())/(events_df %>% filter(classification == 'public_research', event_type %in% c('priv_hs','pub_hs'),univ_state != event_state) %>% count() %>% as.numeric())*100, digits=1)`% for public research universities.

Even though selective private universities tend to emphasize visits to private high schools more than public research universities, both groups visited private high schools at a much higher rate than public high schools. Figure \@ref(fig:actual-proportional-pubu) (public universities) and Figure \@ref(fig:actual-proportional-privu) (private universities) show the actual number of private high school visits made compared to the hypothetical number of private high school visits if each public and private high school in the nation had an equal probability of receiving a visit.^[When calculating the hypothetical proportional number of visits for each university, we only consider high schools in states that received at least one visit to a public or private high school from that university. Our rationale for this decision is that it is unhelpful to include schools in states that the university ignored entirely.] Figure \@ref(fig:actual-proportional-pubu) shows that the actual number of private high school visits exceeded the proportional number of private high school visits for 13 of the 15 public research universities in our sample, the exceptions being UC-Riverside and SUNY Stony Brook. For UC-Berkeley, private school visits represented 27% of all high school visits but would have represented 19% had UC-Berkeley visited public and private high schools at the same rate. For the University of Georgia, which represented the extreme for public universities, private high school visits represented 37% of all high school visits but would have represented 19% had the university visited public and private high schools at the same rate. Figure \@ref(fig:actual-proportional-privu) shows that this disparity between actual versus proportional visits was much greater for private universities than public universities. Catholic universities made the most disproportionate number of private high school visits. For Notre Dame, private high school visits represented 63% of actual high school visits but would have represented 18% of proportional visits.

@@

Research question 2 asks, "To what extent do public research universities and selective private universities visit overlapping sets of private high schools?" Answering this question requires social network analysis techniques. As described in the methods section, we began by constructing a two-mode social network of visits to private high schools (mode 1) by universities (mode 2). Each university and each private high school is a "vertex." A visit from a university to a high school is an "edge." Because large two-mode networks like our school-university network are difficult to visualize, two-mode networks are often analyzed as one-mode networks [@RN2457;@RN4663]. We transformed our two-mode school-university network into a one-mode university network in which each university is a vertex and two universities share an edge if both universities visited at least one high school in common. The *weight* of an edge indicates the number of high schools that both universities visited. For example, Appendix Table \@ref(tab:pubu-count-matrix) shows the matrix of weights for the set of public research universities. We created three separate one-mode university networks: a public university network; a private university network; a public and private university network.

@@

Table \@ref(tab:pubu-percent-matrix) provides a tabular representation of the extent to which public universities visited overlapping sets of private high schools. For each public university (column), each cell of Table \@ref(tab:pubu-percent-matrix) shows the number of visited high schools in common with the row university as a percent of the total number of high schools visited by the column university. This indicates the extent to which public universities were competing for students at the same set of private schools. For example, column 2 shows that of the 396 different private high schools visited by the University of South Carolina, the University of Alabama visited 71.7% of these high schools and CU Boulder visited 39.1% of these high schools. Appendix Table \@ref(tab:pubu-count-matrix) shows counts rather than percentages.

@@

Next, we examine overlap in a network that contains the 15 public universities and the 14 private universities. Table \@ref(tab:privu-percent-visited-by-pubu-matrix) shows, for each private university, the percentage of high schools that were also visited by each public university. This table shows the extent to which public universities were competing at private high schools targeted by private universities. For example, of the 625 private high schools visited by Notre Dame, the University of Alabama visited 52.2% of these schools. Similarly, CU Boulder visited 46.4% of the 435 schools that TCU visited and 42.4% of the 377 schools that Northwestern visited. The University of Arkansas competed heavily at schools visited by Baylor; Rutgers and UMass-Amherst competed heavily at schools visited by Stevens Institute of Technology; and UC-Berkeley competed at schools visited by Northwestern, Tufts, and Emory.

@@

Universities located near the center of Figure \@ref(fig:plot-1mode-u) tended to be universities with the highest degree centrality (i.e., those that visited the most schools). The top nine universities in terms of degree centrality are (number of unique private high schools visited in parentheses): University of Alabama (759), Notre Dame (625), Villanova (563), SMU (550), TCU (435), Tulane (430), University of South Carolina (396), Northwestern (377), and CU Boulder (362). Notably, while the majority of universities located near the center of Figure \@ref(fig:plot-1mode-u) are private, three are public. The most peripheral universities -- UC-Riverside,  SUNY Stony Brook, and University of Nebraska-Lincoln -- are all public.
